A Large-scale Study of Polymorphism in R

### 2019-03-17 paper/main.tex
talked about genthat, and working on method

%
\section{The R Programming Language}

@@

Interestingly, we are somewhat able to replicate this using only package
code.  CRAN, for instance, has a policy that packages made available on the
CRAN platform should be accompanied by a series of examples, tests, and/or
vignettes showing off how the code is intended to be used.  Nominally
unrelated, the {\tt genthat} project aims to trace this swath of example
code to generate new examples which package designers may not have thought
of.  Crucially, {\tt genthat} outputs complete trace information for each of
the functions that it runs so that they can be run again in a fresh R
instance (e.g., by capturing the random seed, environment variables, etc.).
We can leverage this trace information to glean type and attribute patterns
of functions, and since these traces correspond to examples of the packages
being used (rather than just arbitrarily running package code), we have a
reasonable approximation of how R programmers would use the packages.

\AT{We use/modify the {\tt genthat} tracer, probably worth going in to more
  detail here.}
%

@@

First, we leverage {\tt genthat}'s tracer to trace function executions.  The
idea here is that we don't really have access to non-library code written in
R, as general use patterns are (possibly?) to write small scripts which
analyze some bit of data and possibly visualize results.  The goal is likely
not to build big working systems, instead to explore data with by writing
and rewriting small scripts, ad infinitum.  In looking at package tests,
vignettes, and examples, we are painting a picture of how the package
designers intended their packages to be used, which we believe is a close
approximation of what R users would do.

{\tt genthat} generates a trace file which can be loaded into R and contains
all of the necessary information to rerun the tested function.  This
includes the random seed and all relevant environment data.  We utilize this
trace data to collect type information: we run each argument in the
specified environment, and collect type information through R's {\tt typeof}
function, and attribute information through R's {\tt attributes} function.

>>>>

%
\section{The R Programming Language}
\label{sec:rlang}

@@

Interestingly, we are somewhat able to replicate this using only package
code.  CRAN, has a policy that packages made available on the
CRAN platform should be accompanied by a series of examples, tests, and/or
vignettes showing off how the code is intended to be used.  Nominally
unrelated, the {\tt genthat} project~\cite{kvrikava2018tests} traces this swath of example
code to generate new examples which package designers may not have thought
of.  As it happens, we can leverage {\tt genthat} to extract dynamic type
information from running package test code, which presumably indicates
how package code is intended to be used.

\AT{Draft}

We modify the {\tt genthat} tracer to capture the type information of arguments.
{\tt genthat} generates traces for each unique function invocation with the express purpose of re-invoking the function at a later time.
Instead of generating traces for these invocations, we can instead capture type information of the arguments and return values of functions.
This is done by evaluating the arguments and return values of functions at the end of the function's scope in the relevant environment.
The sort of types and granularity of the type information that are generated can be specified in our tracer:
Essentially, our tracer is parameterized over a user-specified type system.
For instance, if desired our tracer can dig into lists and data frames to get the types of list elements and fields.
Our tracer produces trace results for each unique type signature encountered, which may then be used for data analysis.

An unfortunate consequence of R's most common usage pattern is the lack of "deployed" R code.
R is most often used for exploratory data analysis, and very little of this analysis code is made publicly available on e.g. GitHub.
The {\tt genthat} angle is thus desirable:
With the scarcity of non-library code, examples of package usage is about as representative as one can hope for of the sort of code people will write in their analysis pipelines.

We can do a little bit better than merely package usage examples, though.
Available through CRAN is a list of {\it reverse dependencies} of packages:
for some package {\tt p}, a list of all packages which require {\tt p} in some way.
Our tracer is able to leverage and trace this code as well.
When running a package, we also trace how all inherited functions are used, thus painting an even clearer picture of function usage.
This additional data is considered alongside similar data collected for all other packages, as well as the data obtained by tracing the package's examples.

% Some prose we might want to pull from.
%First, we leverage {\tt genthat}'s tracer to trace function executions.  The
%idea here is that we don't really have access to non-library code written in
%R, as general use patterns are (possibly?) to write small scripts which
%analyze some bit of data and possibly visualize results.  The goal is likely
%not to build big working systems, instead to explore data with by writing
%and rewriting small scripts, ad infinitum.  In looking at package tests,
%vignettes, and examples, we are painting a picture of how the package
%designers intended their packages to be used, which we believe is a close
%approximation of what R users would do.

@@

%
%
%
%
\subsection{Data Collection Pipeline}

As we mentioned in the previous section, we modify the {\tt genthat} tracer to collect and output the type information for all function invocations.

\AT{somewhere}
Our analysis can be parameterized by a user-specified ``type system'', simply a map between existing dynamically observable type information and user-defined types.
If no type system is specified, the analysis gains as much information as it can.
For instance, if the analysis encounters {\tt typeof(x) == "double"}, it will look at the length of {\tt x} to determine if {\tt x} is a scalar or a vector, generating the annotation {\tt scalar/double} or {\tt vector/double} accordingly.
Note that this is imperfect, as a vector of unit length would (perhaps incorrectly) show up as a {\tt scalar/X}, but any other invocation where the vector does not have unit length will generate the correct {\tt vector/X} type, and the distinction between scalars and vectors can be later collapsed if desired.
Note also that the type system described here is the \AT{SOMETHING} Type System, described in detail in Section~\AT{TODO}.

The process for tracing one package {\tt p} is as follows:

\begin{enumerate}

	\item first, all functions in {\tt p} are instrumented with on-exit hooks.
	These hooks call a number of R's available reflection functions on each of the arguments, as well as the return type:
	The called functions are {\tt typeof}, {\tt class}, and {\tt attributes}.
	See Section~\ref{sec:rlang} for details on these functions.

	\item then, depending on the specified type

\end{enumerate}
